treats manual

### 2024-08-28 inst/gitbook/02_make.traits.Rmd
pre-update

---
title: "`make.traits`"
author: "Thomas Guillerme (guillert@tcd.ie)"
date: "`r Sys.Date()`"
bibliography: [../references.bib]
output:
  html_document: default
  pdf_document: default
---

@@

## Mapping traits on a tree: `map.traits`

>>>>

---
title: "treats manual"
author: "Thomas Guillerme (t.guillerme@sheffield.ac.uk)"
date: "`r Sys.Date()`"
bibliography: [../References.bib]
output:
  html_document: default
  pdf_document: default
---

@@

## Mapping traits on a tree: `map.traits` {#map.traits}

### 2024-08-28 inst/gitbook/06_plotting_treats_objects.Rmd

---
title: "`plot.treats`"
author: "Thomas Guillerme (guillert@tcd.ie)"
date: "`r Sys.Date()`"
# bibliography: [../References.bib, ../packages.bib]
output:
  html_document: default
  pdf_document: default
---

@@

```{r, fig.height = 6, fig.width = 6}
## Specifying a 3D trait process
my_3D_trait <- make.traits(n = 3, )
## Simulating a birth-death tree with that 3D trait
my_data <- treats(bd.params = list(extinction = 0.2),
                  stop.rule = list(max.living = 50),
                  traits    = my_3D_trait)
```

@@

If you like the plotting style of `treats` trees and data, you can always use the `make.traits` function to plot a tree and some traits as a `"treats"` object.

### Discrete plots

### Continuous plots (with [mapping](#map.traits))

>>>>

---
title: "treats manual"
author: "Thomas Guillerme (t.guillerme@sheffield.ac.uk)"
date: "`r Sys.Date()`"
bibliography: [../References.bib]
output:
  html_document: default
  pdf_document: default
---

@@

```{r, fig.height = 6, fig.width = 6}
## Specifying a 3D trait process
my_3D_trait <- make.traits(n = 3)
## Simulating a birth-death tree with that 3D trait
my_data <- treats(bd.params  = list(extinction = 0.2),
                  stop.rule  = list(max.living = 50),
                  traits     = my_3D_trait,
                  null.error = 100)
```

@@

If you like the plotting style of `treats` trees and data, you can always use the `make.treats` function to plot a tree and some traits as a `"treats"` object.
If you have a dataset that contains trait data for tips and nodes you can transform it into a `"treats"` object and plot it using the generic `plot.treats` function:

```{r}
set.seed(1)
## Generating some random tree
my_tree <- makeNodeLabel(rcoal(20))

## Generating some random discrete data
discrete_data <- sample(c(1,2), 39, replace = TRUE)
names(discrete_data) <- c(my_tree$tip.label, my_tree$node.label)

## Creating a treats object with this data
discrete_treats <- make.treats(tree = my_tree, data = discrete_data)

## Plot the data together
plot(discrete_treats, legend = TRUE)
```

The same goes for continuous data:

```{r}
## Generating some random continuous data
continuous_data <- rnorm(39)
names(continuous_data) <- c(my_tree$tip.label, my_tree$node.label)

## Creating a treats object with this data
continuous_treats <- make.treats(tree = my_tree, data = continuous_data)

## Plot the data together
plot(continuous_treats)
```
